Another Look at Regularized Discriminant Analysis for High-Dimensional Classification

### 2013-04-21 reports/paper-rda.Rnw
Updated paper to match the majority of Biometrika's requirements. Still remaining: tweak notation and update figures.

@@

# Summary of the classification error rates for each data set and number of
# variables selected, d.
error_rates <- subset(results, select = -c(lambda, gamma))
m_error <- melt(error_rates, variable.name = "Classifier", value.name = "Error")
error_rates <- ddply(m_error, .(data_set, d, Classifier), plyr:::summarize,
                     Avg_Error = mean(Error), Num_Reps = length(Error))
error_rates <- ddply(error_rates, .(data_set, d, Classifier), transform,
                     SE_Error = sqrt(Avg_Error * (1 - Avg_Error) / Num_Reps))
error_rates <- subset(error_rates, select = -Num_Reps)

@@

A reliable supervised classification of tumors from gene expression microarray
data is essential for successful diagnosis and treatment of diseases such as
cancer; however, the analysis of microarray data often presents difficult
challenges because the number of gene expression levels is typically much
greater than the sample size. Consequently, several sparse and regularized
classifiers have been proposed, often with strict covariance assumptions or
eigenvalue adjustments to ensure that the classifiers are well-posed. Here, we
consider an improved version of the regularized discriminant analysis
(\emph{RDA}) classifier popularized by \cite{Friedman:1989tm}. That is, we
propose the high-dimensional \emph{RDA} (\emph{HDRDA}) classifier, a novel
modification of the \emph{RDA} classifier, that provides a straightforward
interpretation and employs a new dimension-reduction technique to enable a
computationally efficient model-selection procedure, thus overcoming the
computational shortcomings of the \emph{RDA} classifier for high-dimensional
data. Moreover, we demonstrate with several high-dimensional microarray data
sets that our proposed classifier is competitive and often superior to several
recently proposed sparse and regularized classifiers in terms of classification
accuracy. Finally, we provide an implementation of the \emph{HDRDA} classifier in
the {\tt sparsediscrim} R package, available on CRAN.

@@

Microarrays have enabled researchers to measure gene-expression levels of
thousands of genes in a single experiment to identify a subset of genes and
their expression patterns that either characterize a particular cell state or
facilitate an accurate prediction of future cell state.  We focus on the
classification of tissues and other biological samples into one of $K$ disjoint,
predefined cancer types, where our goal is to ascertain the class membership of
a new tissue sample based on gene-expression microarray data. However,
microarray data typically consist of a large number of gene expressions $p$
relative to the total sample size $N$. Moreover, standard benchmark microarray
data sets usually consist of between $p = 10,000$ and $55,000$ probes and fewer
than $N = 100$ observations, in which case standard supervised-learning methods
exhibit poor classification performance because of the \emph{curse of
  dimensionality} \citep{Bellman:1961tn}. Furthermore, well-known classifiers,
including linear discriminant analysis \emph{(LDA)} and quadratic discriminant
analysis \emph{(QDA)}, are incalculable when $p > N$ because the sample
covariance matrices employed in both classifiers are singular. As a result, many
researchers, including \cite*{Mai:2012bf}, \cite*{Fan:2012iq},
\cite*{Merchante:2012vk}, \cite*{Clemmensen:2011kr}, and \cite*{Huang:2010ju},
have introduced restrictions on the population covariance matrices in the
\emph{LDA} and \emph{QDA} models to reduce number of covariance parameters that
must be estimated to ensure that the resulting classifiers are calculable.  For
instance, \cite{Dudoit:2002ev} have proposed the diagonal \emph{LDA}
(\emph{DLDA}) classifier, a modification of the \emph{LDA} classifier where the
class covariance matrices are assumed to be diagonal.

Alternatively, several researchers, including \cite*{Zhang:2010va},
\cite{Ye:2009gd}, \cite{Ji:2008wp}, \cite*{Guo:2007te},
\cite*{Srivastava:2007ud}, \cite{Ye:2006vx}, and \cite{Ye:2005uu}, have proposed
regularization methods to improve the estimation of the population covariance
matrices before applying the \emph{LDA} and \emph{QDA}
classifiers. Covariance-matrix regularization methods often are based on the
well-known ridge-regression approach of \cite{Hoerl:1970cd} and adjust the
eigenvalues of the sample covariance matrices to ensure positive
definiteness. For excellent overviews of the regularization problem in
discriminant analysis, see \cite{Murphy:2012uq} and \cite{Mkhadri:1997gy}.

Here, we propose a generalization of the popular \emph{RDA} classifier of
\cite{Friedman:1989tm}, who has proposed pooling of the covariance-matrix
estimators from the \emph{LDA} and \emph{QDA} classifiers and has additionally
incorporated a covariance-matrix shrinkage term to ensure that the \emph{RDA}
classifier is well-posed. \cite{Friedman:1989tm} has shown that the \emph{RDA}
classifier can attain improved classification accuracy, compared to the
\emph{LDA} and \emph{QDA} classifiers when $p < N$. However, in its original
form, the \emph{RDA} classifier is impractical when $p \gg N$ because the
\emph{RDA} decision rule requires the calculation of the inverses and
determinants of the $p \times p$ covariance-matrix estimators for each
class. Furthermore, the \emph{RDA} classifier has two tuning parameters that are
selected via cross-validation, which adds a substantial computational burden to
the model-selection process. The \emph{RDA} classifier is ill-advised on modern
data sets because the model selection requires $O(Kp^3)$ calculations for each
tuning-parameter pair.

In this paper we modify the \emph{RDA} classifier by first simplifying the
pooling matrix employed in the \emph{RDA} classifier. We construct a convex
combination of the covariance-matrix estimators employed in the \emph{LDA} and
\emph{QDA} classifiers and demonstrate that this simplification improves the
interpretation of the pooling operation of the \emph{RDA} classifier.
\cite{Clemmensen:2011kr} have argued that the assumption of equal population
covariance matrices within the \emph{LDA} classifier is often too stringent in
practice. The \emph{HDRDA} classifier allows us to relax the linearity
assumption by including unequal covariance matrices in the model. The
\emph{HDRDA} classifier is then a compromise between the \emph{LDA} and
\emph{QDA} classifiers, so that the \emph{HDRDA} classifier includes both the
\emph{LDA} and \emph{QDA} classifiers as special cases. We then establish the
\emph{HDRDA} classifier by applying covariance-matrix regularization with a
parameterization that includes a large family of shrinkage methods, such as the
methods from \cite{Srivastava:2007ww}, \cite{Rao:1971ul}, and a wide variety of
other shrinkage methods studied by \cite{Ramey:2013ji} and \cite{Xu:2009fl}.

We show that the \emph{HDRDA} classifier is invariant to any adjustments of the
eigenvalues corresponding to the null space of the pooled sample covariance
matrix. Applying a reasoning similar to \cite{Ye:2006tq}, we decompose the
\emph{RDA} classifier and show that the sum related to the null space of the
pooled sample covariance matrix can be discarded. Hence, no gain in
classification performance can be obtained by adjusting these zero
eigenvalues. This invariance property addresses the computational shortcomings
of the \emph{RDA} classifier and provides a substantial increase in
computational efficiency and model selection. In fact, we show that the inverse
and determinants can be reduced to calculations of $n_k \times n_k$ matrices,
where $n_k$ is the sample size of the $k$th class, $k = 1, \ldots, K$. This
reduction in calculation is clearly advantageous because $n_k$ is often quite
small, say, 15--20, in microarray data sets.

We also demonstrate that our \emph{HDRDA} classifier is competitive with and
often superior to several recent classifiers designed for small-sample,
high-dimensional data. In particular, we compare our \emph{HDRDA} classifier
with the \emph{DLDA} classifier from \cite{Dudoit:2002ev} and the recently
proposed classifiers from \cite{Guo:2007te}, \cite{Pang:2009ik},
\cite{Tong:2012hw}, and \cite{Witten:2011kc}. We conduct our comparison of the
classification accuracy of the competing classifiers using four small-sample,
high-dimensional microarray data sets.

We have organized the remainder of the paper as follows. In Section 2 we review
the \emph{RDA} classifier from \cite{Friedman:1989tm} and discuss our
generalization in Section 3. In Section 4 we describe four small-sample,
high-dimensional microarray data sets and then demonstrate that our proposed
\emph{HDRDA} classifier is often superior to the competing classifiers in terms
of classification accuracy. We then conclude with a brief discussion in Section
5.

@@

In discriminant analysis we wish to assign correctly an unlabeled
$p$-dimensional observation vector $\bm x$ to one of $K$ unique, known classes
(or populations) by constructing a classifier from $n$ training observations
that can accurately predict the class membership of $\bm x$. Let $\bm x_i =
(x_{i1}, \ldots, x_{ip}) \in \mathbb{R}_{p \times 1}$ be the $i$th observation
$(i = 1, \ldots, N)$ with true, unique membership $y_i \in \{\omega_1, \ldots,
\omega_K\}$, where $\mathbb{R}_{a \times b}$ denotes the matrix space of all $a
\times b$ matrices over the real field $\mathbb{R}$. We assume that $(\bm x_i,
y_i)$ is a realization from a mixture distribution $p(\bm x) = \sum_{k=1}^K
p(\bm x | \omega_k) p(\omega_k)$, where $p(\bm x | \omega_k)$ is the probability
density function (PDF) of the $k$th class, $p(\omega_k)$ is prior probability of
class membership of the $k$th class.

The \emph{QDA} classifier is the optimal Bayesian decision rule with respect to
a $0-1$ loss function when $p(\bm x | \omega_k)$ is the PDF of the multivariate
normal distribution with known mean vector $\bm\mu_k \in \mathbb{R}_{p \times
  1}$ and known covariance matrix $\bm\Sigma_k \in \mathbb{R}_{p \times p}^{>}$,
$k = 1, 2, \ldots, K$, where $\mathbb{R}_{p \times p}^{>}$ denotes the cone of
real $p \times p$ positive-definite matrices. Because the parameters are
typically unknown, we estimate the unknown parameters $\bm \mu_k$ and
$\bm\Sigma_k$ with their maximum likelihood estimators (MLEs) and substitute the
MLEs into the \emph{QDA} classifier. Assuming that the prior probabilities of
class membership $p(\omega_k)$ are equal for $k = 1, \ldots, K$, we assign an
unlabeled observation $\bm x$ to class $\omega_k$ with the sample \emph{QDA}
classifier
\begin{align}
	D_{QDA}(\bm x) = \argmin_{k}  (\bm x - \xbar_k)'\widehat{\bm\Sigma}_k^{-1}(\bm x - \xbar_k)  + \ln |\widehat{\bm\Sigma}_k|, \label{eq:qda}
\end{align}
where $\xbar_k$ and $\widehat{\bm\Sigma}_k$ are the MLEs for $\bm \mu_k$ and
$\bm \Sigma_k$, respectively. If we assume further that the population
covariance matrices are equal for each class (i.e., $\bm\Sigma_k = \bm\Sigma$
for all $k$), then the pooled sample covariance matrix $\widehat{\bm\Sigma}$ is
substituted for $\widehat{\bm \Sigma}_k$ in \eqref{eq:qda}, where
\begin{align}
	\widehat{\bm\Sigma} = \frac{1}{N} \sum_{k=1}^K n_k \widehat{\bm\Sigma}_k. \label{eq:pooled-cov}
\end{align}
is the MLE for $\bm \Sigma$. In this case \eqref{eq:qda} reduces to Fisher's
\emph{LDA} classifier, and the log-determinant is omitted because it is constant
across all classes.

We also require the following additional notation throughout the document. Let
$\bm A \oplus \bm B$ denote the direct sum of $\bm A \in \mathbb{R}_{r \times
  r}$ and $\bm B \in \mathbb{R}_{s \times s}$ \citep[Chapter
  1]{Lutkepohl:1996uz}. Let $\bm A^+$ and $\mathcal{N}(\bm A) = \{\bm z \in
\mathbb{R}_{p \times 1} : \bm A \bm z = \bm 0\}$ denote the Moore-Penrose
pseudoinverse and null space of $\bm A \in \mathbb{R}_{m \times p}$,
respectively. Let $V^{\perp} = \{\bm y \in \mathbb{R}_{p \times 1} : \bm y'\bm v
= 0\ \forall \bm v \in V \}$ denote the orthogonal complement of a vector space
$V \subset \mathbb{R}_{p \times 1}$. Let $\bm I_m$ be the $m \times m$ identity
matrix, and let $\bm 0_{m \times p}$ be the $m \times p$ matrix of zeros, such
that $\bm 0_m$ is understood to denote $\bm 0_{m \times m}$. Lastly, for $c \in
\mathbb{R}$, let $c^+ = 1/c$ if $c \ne 0$ and $0$ otherwise.

@@

Several regularization methods, such as the methods considered by
\citet*{Guo:2007te}, \cite{Mkhadri:1995jp}, and \citet*{Xu:2009fl}, have been
proposed in the literature to adjust the eigenvalues of $\widehat{\bm \Sigma}_k$
so that \eqref{eq:qda} is calculable and has reduced variability in the standard
bias-variance tradeoff.  A common form of the covariance-matrix regularization
applies a shrinkage factor $\gamma_k > 0$, so that
\begin{align}
	\widehat{\bm \Sigma}_k(\gamma_k) =  \widehat{\bm \Sigma}_k + \gamma_k \bm I_p, \label{eq:ridge-estimator}
\end{align}
similar to a method employed in ridge regression \citep{Hoerl:1970cd}. Equation
\eqref{eq:ridge-estimator} effectively \emph{shrinks} the sample covariance
matrix $\widehat{\bm\Sigma}_k$ towards $\bm I_p$, thereby increasing the
eigenvalues of $\widehat{\bm\Sigma}_k$ by $\gamma_k$. Specifically, the zero
eigenvalues are replaced with $\gamma_k$, so that \eqref{eq:ridge-estimator} is
positive definite. For additional covariance-matrix regularization methods, see
\cite{Ramey:2013ji}, \cite{Xu:2009fl}, and \cite{Ye:2009gd}.

@@

\cite{Friedman:1989tm} has proposed the \emph{RDA} classifier by incorporating a
weighted average of the sample covariance matrix $\widehat{\bm \Sigma}_k$ for
class $\omega_k$ and the pooled sample covariance matrix $\widehat{\bm\Sigma}$
to estimate the covariance matrix for class $\omega_k$ with
\begin{align}
  \widehat{\bm\Sigma}_k(\lambda) = n_k^{-1}(\lambda) \bm S_k(\lambda),\label{eq:sig-lambda}
\end{align}
where $\lambda \in [0, 1]$, $n_k(\lambda) = (1 - \lambda) n_k + \lambda N$, $\bm
S_k = n_k \widehat{\bm\Sigma}_k$, $\bm S = \sum_{k=1}^K \bm S_k$, and $\bm
S_k(\lambda) = (1 - \lambda) \bm S_k + \lambda \bm S$. We can interpret
\eqref{eq:sig-lambda} as a covariance matrix estimator for class $\omega_k$ that
borrows from $\widehat{\bm\Sigma}$ in \eqref{eq:pooled-cov} to better estimate
$\bm \Sigma_k$. If $\lambda = 0$, \eqref{eq:sig-lambda} is equal to
$\widehat{\bm\Sigma}_k$ used in the \emph{QDA} classifier in
\eqref{eq:qda}. Alternatively, if $\lambda = 1$, \eqref{eq:sig-lambda} reduces
to \eqref{eq:pooled-cov} used in the \emph{LDA} classifier.

To further improve the estimation of $\bm \Sigma_k$ and to stabilize the inverse
of \eqref{eq:sig-lambda}, \cite{Friedman:1989tm} has proposed the biased
\emph{RDA} covariance-matrix estimator
\begin{align}
	\widehat{\bm\Sigma}_k(\lambda, \gamma) = (1 - \gamma) \widehat{\bm\Sigma}_k(\lambda) + \gamma \frac{\tr\{\widehat{\bm\Sigma}_k(\lambda)\}}{p} \bm I_p,\label{eq:sig-rda}
\end{align}
for class $\omega_k$, where $\gamma \in [0, 1]$ is a regularization parameter
that controls the shrinkage of \eqref{eq:sig-rda} towards $\bm I_p$, which is
weighted by the the average of the eigenvalues of \eqref{eq:sig-lambda}. Thus,
the \emph{pooling} parameter $\lambda$ controls the amount that we borrow from
$\widehat{\bm\Sigma}$ to estimate $\bm \Sigma_k$, and the \emph{shrinkage}
parameter $\gamma$ determines the amount of shrinkage applied.

\cite{Friedman:1989tm} substituted \eqref{eq:sig-rda} into \eqref{eq:qda} to
obtain the \emph{RDA} classifier
\begin{align}
	D_{RDA}(\bm x) = \argmin_{k}  (\bm x - \xbar_k)'\widehat{\bm\Sigma}_k(\lambda, \gamma)^{-1}(\bm x - \xbar_k)  + \ln |\widehat{\bm\Sigma}_k(\lambda, \gamma)|. \label{eq:rda}
\end{align}

\section{The HDRDA Classifier}

>>>>

# To abide by Biometrika's requirements of no abbreviations, we rename the
# results of our HDRDA classifier as simply "Ours."
results$Ours <- results$HDRDA
results <- subset(results, select = -HDRDA)

@@

# Summary of the classification error rates for each data set and number of
# variables selected, d.
error_rates <- subset(results, select = -c(lambda, gamma))
m_error <- melt(error_rates, variable.name = "Classifier", value.name = "Error")
m_error$Classifier <- factor(m_error$Classifier)
error_rates <- ddply(m_error, .(data_set, d, Classifier), plyr:::summarize,
                     Avg_Error = mean(Error), Num_Reps = length(Error))
error_rates <- ddply(error_rates, .(data_set, d, Classifier), transform,
                     SE_Error = sqrt(Avg_Error * (1 - Avg_Error) / Num_Reps))
error_rates <- subset(error_rates, select = -Num_Reps)

@@

A reliable supervised classification of tumors from gene expression microarray
data is essential for successful diagnosis and treatment of diseases such as
cancer; however, the analysis of microarray data often presents difficult
challenges because the number of gene expression levels is typically much
greater than the sample size. Consequently, several sparse and regularized
classifiers have been proposed, often with strict covariance assumptions or
eigenvalue adjustments to ensure that the classifiers are well-posed. Here, we
propose a novel modification of the regularized discriminant analysis classifier
popularized by \cite{Friedman:1989tm}. Our proposed classifier provides a
straightforward interpretation and employs a new dimension-reduction technique
to enable a computationally efficient model-selection procedure, thus overcoming
the computational shortcomings of \citeauthor{Friedman:1989tm}'s classifier for
high-dimensional data. Moreover, we demonstrate with several high-dimensional
microarray data sets that our proposed classifier is competitive and often
superior to several recently proposed sparse and regularized classifiers in
terms of classification accuracy. Finally, we provide an implementation of our
proposed classifier in the {\tt sparsediscrim} R package, available on CRAN.

@@

Microarrays have enabled researchers to measure gene-expression levels of
thousands of genes in a single experiment to identify a subset of genes and
their expression patterns that either characterize a particular cell state or
facilitate an accurate prediction of future cell state.  We focus on the
classification of tissues and other biological samples into one of $K$ disjoint,
predefined cancer types, where our goal is to ascertain the class membership of
a new tissue sample based on gene-expression microarray data. However,
microarray data typically consist of a large number of gene expressions $p$
relative to the total sample size $N$. Moreover, standard benchmark microarray
data sets usually consist of between $p = 10,000$ and $55,000$ probes and fewer
than $N = 100$ observations, in which case standard supervised-learning methods
exhibit poor classification performance because of the \emph{curse of
  dimensionality} \citep{Bellman:1961tn}. Furthermore, some well-known
classifiers, such as linear and quadratic discriminant analysis, are
incalculable when $p > N$ because the sample covariance matrices employed are
singular. As a result, many researchers, including \cite{Mai:2012bf},
\cite{Fan:2012iq}, \cite{Merchante:2012vk}, \cite{Clemmensen:2011kr}, and
\cite{Huang:2010ju}, have introduced restrictions on the population covariance
matrices in linear and quadratic discriminant analysis to reduce number of
covariance parameters that must be estimated to ensure that the resulting
classifiers are calculable. For instance, \cite{Dudoit:2002ev} have modified the
linear and quadratic discriminant analysis classifiers by assuming that the
population covariance matrices are diagonal, resulting in classifiers that are
special cases of the Naive Bayes classifier \citep[Chapter 4]{Hastie:2008dt}.

Alternatively, several researchers, including \cite*{Zhang:2010va},
\cite{Ye:2009gd}, \cite{Ji:2008wp}, \cite*{Guo:2007te},
\cite*{Srivastava:2007ud}, \cite{Ye:2006vx}, and \cite{Ye:2005uu}, have proposed
regularization methods to improve the estimation of the population covariance
matrices before applying either linear or quadratic discriminant
analysis. Covariance-matrix regularization methods often are based on the
well-known ridge-regression approach of \cite{Hoerl:1970cd} and adjust the
eigenvalues of the sample covariance matrices to ensure positive
definiteness. For excellent overviews of the regularization problem in
discriminant analysis, see \cite{Murphy:2012uq} and \cite{Mkhadri:1997gy}.

Here, we propose an improvement to the popular regularized discriminant
analysis classifier of \cite{Friedman:1989tm}, who has proposed pooling of the
covariance-matrix estimators from linear and quadratic discriminant analysis and
has additionally incorporated a covariance-matrix shrinkage term to ensure that
the classifier is well-posed. \citeauthor{Friedman:1989tm}'s classifier can
attain improved classification accuracy, compared to linear and quadratic
discriminant analysis when $p < N$. However, in its original form,
\citeauthor{Friedman:1989tm}'s classifier is impractical when $p \gg N$ because
the decision rule requires the calculation of the inverses and determinants of
large $p \times p$ covariance-matrix estimators for each class. Furthermore,
\citeauthor{Friedman:1989tm}'s classifier has two tuning parameters that are
selected via cross-validation, which adds a substantial computational burden to
the model-selection process. \citeauthor{Friedman:1989tm}'s classifier is
ill-advised on modern data sets because the model selection requires $O(Kp^3)$
calculations for each tuning-parameter pair.

In this paper we modify \citeauthor{Friedman:1989tm}'s classifier by first
constructing the pooling matrix as a convex combination of the covariance-matrix
estimators employed in linear and quadratic discriminant analysis and
demonstrate that this simplification improves the interpretation of the pooling
operation of \citeauthor{Friedman:1989tm}'s classifier. Given that the
assumption of equal population covariance matrices within linear discriminant
analysis is often too stringent in practice \citep{Clemmensen:2011kr}, our
proposed classifier can relax the linearity assumption by instead incorporating
unequal covariance matrices while retaining linear and quadratic discriminant
analysis as special cases. We then establish our classifier by applying
covariance-matrix regularization with a parameterization that includes a large
family of shrinkage methods, such as those from \cite{Srivastava:2007ww},
\cite{Rao:1971ul}, and other methods studied by \cite{Ramey:2013ji} and
\cite{Xu:2009fl}.

We show that our proposed classifier is invariant to any adjustments of the
eigenvalues corresponding to the null space of the pooled sample covariance
matrix. Applying a reasoning similar to \cite{Ye:2006tq}, we decompose the
decision rule of our classifier and show that the sum related to the null space
of the pooled sample covariance matrix can be discarded. Hence, no gain in
classification performance can be obtained by adjusting these zero
eigenvalues. This invariance property addresses the computational shortcomings
of \citeauthor{Friedman:1989tm}'s classifier and provides a substantial increase
in computational efficiency and model selection. In fact, we show that the
inverse and determinants can be reduced to calculations of $n_k \times n_k$
matrices, where $n_k$ is the sample size of the $k$th class $(k = 1, \ldots,
K)$. This reduction in calculation is clearly advantageous because $n_k$ is
often quite small in microarray data sets (e.g., 15--20 observations).

We also demonstrate that our proposed classifier is competitive with and often
superior to several recent classifiers designed for small-sample,
high-dimensional data. In particular, we compare our classifier with diagonal
linear discriminant analysis from \cite{Dudoit:2002ev} and the recently proposed
classifiers from \cite{Guo:2007te}, \cite{Pang:2009ik}, \cite{Tong:2012hw}, and
\cite{Witten:2011kc}. We conduct our comparison of the classification accuracy
of the competing classifiers using four small-sample, high-dimensional
benchmark microarray data sets.

@@

\subsection{Preliminaries}
In discriminant analysis we wish to assign correctly an unlabeled
$p$-dimensional observation vector $\bm x$ to one of $K$ unique, known classes
(or populations) by constructing a classifier from $n$ training observations
that can accurately predict the class membership of $\bm x$. Let $\bm x_i =
(x_{i1}, \ldots, x_{ip}) \in \mathbb{R}_{p \times 1}$ be the $i$th observation
$(i = 1, \ldots, N)$ with true, unique membership $y_i \in \{\omega_1, \ldots,
\omega_K\}$, where $\mathbb{R}_{a \times b}$ denotes the matrix space of all $a
\times b$ matrices over the real field $\mathbb{R}$. We assume that $(\bm x_i,
y_i)$ is a realization from a mixture distribution $p(\bm x) = \sum_{k=1}^K
p(\bm x | \omega_k) p(\omega_k)$, where $p(\bm x | \omega_k)$ is the probability
density function of the $k$th class and $p(\omega_k)$ is the prior probability
of class membership of the $k$th class. We further assume that each of
$p(\omega_k)$ is equal $(k = 1, \ldots, K)$.

Quadratic discriminant analysis is the optimal Bayesian decision rule with
respect to a $0-1$ loss function when $p(\bm x | \omega_k)$ is the probability
density function of the multivariate normal distribution with known mean vector
$\bm\mu_k \in \mathbb{R}_{p \times 1}$ and known covariance matrix $\bm\Sigma_k
\in \mathbb{R}_{p \times p}^{>}$, $k = 1, 2, \ldots, K$, where $\mathbb{R}_{p
  \times p}^{>}$ denotes the cone of real $p \times p$ positive-definite
matrices. Because $\bm \mu_k$ and $\bm\Sigma_k$ are typically unknown, we assign
an unlabeled observation $\bm x$ to class $\omega_k$ with the sample quadratic
discriminant analysis classifier
\begin{align}
	D_{QDA}(\bm x) = \argmin_{k}  (\bm x - \xbar_k)'\widehat{\bm\Sigma}_k^{-1}(\bm x - \xbar_k)  + \log |\widehat{\bm\Sigma}_k|, \label{eq:qda}
\end{align}
where $\xbar_k$ and $\widehat{\bm\Sigma}_k$ are the maximum likelihood
estimators for $\bm \mu_k$ and $\bm \Sigma_k$, respectively. If we assume
further that $\bm\Sigma_k = \bm\Sigma$ $(k = 1, \ldots, K)$, then the pooled
sample covariance matrix $\widehat{\bm\Sigma}$ is substituted for $\widehat{\bm
  \Sigma}_k$ in \eqref{eq:qda}, where
\begin{align}
	\widehat{\bm\Sigma} = N^{-1} \sum_{k=1}^K n_k \widehat{\bm\Sigma}_k \label{eq:pooled-cov}
\end{align}
is the maximum likelihood estimator for $\bm \Sigma$. In this case
\eqref{eq:qda} reduces to the sample linear discriminant analysis classifier,
and the log-determinant is omitted because it is constant across all classes.

Throughout the document we also require the following additional notation. Let
$\bm A \oplus \bm B$ denote the direct sum of $\bm A \in \mathbb{R}_{r \times
  r}$ and $\bm B \in \mathbb{R}_{s \times s}$ \citep[Chapter
  1]{Lutkepohl:1996uz}. Let $\bm A^+$ and $\mathcal{N}(\bm A) = \{\bm z \in
\mathbb{R}_{p \times 1} : \bm A \bm z = \bm 0\}$ denote the Moore-Penrose
pseudoinverse and null space of $\bm A \in \mathbb{R}_{m \times p}$,
respectively. Let $V^{\perp} = \{\bm y \in \mathbb{R}_{p \times 1} : \bm y'\bm v
= 0\ \forall \bm v \in V \}$ denote the orthogonal complement of a vector space
$V \subset \mathbb{R}_{p \times 1}$. Let $\bm I_m$ be the $m \times m$ identity
matrix, and let $\bm 0_{m \times p}$ be the $m \times p$ matrix of zeros, such
that $\bm 0_m$ is understood to denote $\bm 0_{m \times m}$. Lastly, for $c \in
\mathbb{R}$, let $c^+ = 1/c$ if $c \ne 0$ and $0$ otherwise.

@@

Several regularization methods, such as the methods considered by
\citet*{Guo:2007te}, \cite{Mkhadri:1995jp}, and \citet*{Xu:2009fl}, have been
proposed in the literature to adjust the eigenvalues of $\widehat{\bm \Sigma}_k$
so that \eqref{eq:qda} is calculable and has reduced variability in the standard
bias-variance tradeoff. A common form of the covariance-matrix regularization
applies a shrinkage factor $\gamma > 0$, so that
\begin{align}
	\widehat{\bm \Sigma}_k(\gamma) =  \widehat{\bm \Sigma}_k + \gamma \bm I_p, \label{eq:ridge-estimator}
\end{align}
similar to a method employed in ridge regression \citep{Hoerl:1970cd}. Equation
\eqref{eq:ridge-estimator} effectively \emph{shrinks} the sample covariance
matrix $\widehat{\bm\Sigma}_k$ towards $\bm I_p$, thereby increasing the
eigenvalues of $\widehat{\bm\Sigma}_k$ by $\gamma$. Specifically, the zero
eigenvalues are replaced with $\gamma$, so that \eqref{eq:ridge-estimator} is
positive definite. For additional covariance-matrix regularization methods, see
\cite{Ramey:2013ji}, \cite{Xu:2009fl}, and \cite{Ye:2009gd}.

@@

\cite{Friedman:1989tm} has proposed regularized discriminant analysis by
incorporating a weighted average of the sample covariance matrix $\widehat{\bm
  \Sigma}_k$ and the pooled sample covariance matrix $\widehat{\bm\Sigma}$ to
estimate $\bm \Sigma_k$ with
\begin{align}
  \widehat{\bm\Sigma}_k(\lambda) = n_k^{-1}(\lambda) \bm S_k(\lambda) \quad (k =
  1, \ldots, K),\label{eq:sig-lambda}
\end{align}
where $\lambda \in [0, 1]$, $n_k(\lambda) = (1 - \lambda) n_k + \lambda N$, $\bm
S_k = n_k \widehat{\bm\Sigma}_k$, $\bm S = \sum_{k=1}^K \bm S_k$, and $\bm
S_k(\lambda) = (1 - \lambda) \bm S_k + \lambda \bm S$. We can interpret
\eqref{eq:sig-lambda} as a covariance-matrix estimator that borrows from
$\widehat{\bm\Sigma}$ in \eqref{eq:pooled-cov} to better estimate $\bm
\Sigma_k$. If $\lambda = 0$, \eqref{eq:sig-lambda} is equal to
$\widehat{\bm\Sigma}_k$ used in \eqref{eq:qda}. Alternatively, if $\lambda = 1$,
\eqref{eq:sig-lambda} reduces to \eqref{eq:pooled-cov}.

To further improve the estimation of $\bm \Sigma_k$ and to stabilize the inverse
of \eqref{eq:sig-lambda}, \cite{Friedman:1989tm} has proposed the biased
covariance-matrix estimator
\begin{align}
	\widehat{\bm\Sigma}_k(\lambda, \gamma) = (1 - \gamma) \widehat{\bm\Sigma}_k(\lambda) + \gamma \frac{\tr\{\widehat{\bm\Sigma}_k(\lambda)\}}{p} \bm I_p,\label{eq:sig-rda}
\end{align}
where $\gamma \in [0, 1]$ is a regularization parameter that controls the
shrinkage of \eqref{eq:sig-rda} towards $\bm I_p$ and is weighted by the average
of the eigenvalues of \eqref{eq:sig-lambda}. Thus, the \emph{pooling} parameter
$\lambda$ controls the amount borrowed from $\widehat{\bm\Sigma}$ to estimate
$\bm \Sigma_k$, and the \emph{shrinkage} parameter $\gamma$ determines the
amount of eigenvalue shrinkage applied. Finally, by substituting
\eqref{eq:sig-rda} into \eqref{eq:qda}, we have the regularized discriminant
analysis classifier
\begin{align}
	D_{RDA}(\bm x) = \argmin_{k}  (\bm x - \xbar_k)'\widehat{\bm\Sigma}_k(\lambda, \gamma)^{-1}(\bm x - \xbar_k)  + \log |\widehat{\bm\Sigma}_k(\lambda, \gamma)|. \label{eq:rda}
\end{align}
If $\gamma$ is sufficiently small, \eqref{eq:sig-rda} can be singular or
near-singular. For this case \cite{Friedman:1989tm} has suggested replacing the
zero eigenvalues of \eqref{eq:sig-rda} with a small positive constant that
permits a numerically stable matrix inversion.

\section{High-Dimensional Regularized Discriminant Analysis}

\subsection{Improvements to \citeauthor{Friedman:1989tm}'s Classifier}
